---
title: "Mapping data in R: `ggmap` tutorial"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r libraries}
library(tidyverse)     # for data cleaning and plotting
library(googlesheets4) # for reading googlesheet data
library(lubridate)     # for date manipulation
library(openintro)     # for the abbr2state() function
library(palmerpenguins)# for Palmer penguin data
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(ggthemes)      # for more themes (including theme_map())
gs4_deauth()           # To not have to authorize each time you knit.
theme_set(theme_minimal())
```

```{r data}
# Lisa's Mallorca cycling data
mallorca_bike_day7 <- read_csv("https://www.dropbox.com/s/zc6jan4ltmjtvy0/mallorca_bike_day7.csv?dl=1") %>% 
  select(1:4, speed)
```


In this demo, we are going to use `ggmap` to plot a bike ride I went on in Mallorca, Spain in 2018. First, let's take a quick look at the data. The data were originally from my Garmin in a GPX file. I used a function called `readGPX()` from the `plotKML` library to get it into R - GPX data are UG-LY!

Next, let's take a look at a map of [Mallorca](http://maps.stamen.com/#terrain/9/39.6078/3.0120) to get an idea of where it is for those who don't know. This is a stamen map, which is the same type of map we'll be bringing into R to make the map.


Now, let's get the map of Mallorca. To decide the `bbox` values, I go to [openstreetmap](https://www.openstreetmap.org), put in my location, set to the right zoom, and click export. Then, I use the numbers in the corresponding boxes. This map also has a zoom number, as did the stamen map in the link above.

```{r}
mallorca_map <- get_stamenmap(
    bbox = c(left = 2.2, bottom = 39.2, right = 3.7, top = 40), 
    maptype = "terrain",
    zoom = 10
)
```


And, let's plot the map to make sure it looks like what we expected:

```{r}
ggmap(mallorca_map)
```

Now, we are going to plot the ride I went on (on the 7th day I was there). Let's start by plotting points. 

```{r}
ggmap(mallorca_map) +
  geom_point(data = mallorca_bike_day7, 
            aes(x = lon, y = lat),
            size = .3) +
  theme_map()
```

I want to make some changes, including:

* Zooming in on the map more (ie. change `bbox` argument)  
* Make the ride a line, rather than a point (use `geom_path()` rather than `geom_line()` to connect the points in order)  
* Color the line by elevation and change the color scale
* Eliminate the background color of the legend so it doesn't overlap the map

```{r}
mallorca_map <- get_stamenmap(
    bbox = c(left = 2.28, bottom = 39.41, right = 3.03, top = 39.8), 
    maptype = "terrain",
    zoom = 11
)
ggmap(mallorca_map) +
  geom_path(data = mallorca_bike_day7, 
             aes(x = lon, y = lat, color = ele),
             size = .5) +
  scale_color_viridis_c(option = "magma") +
  theme_map() +
  theme(legend.background = element_blank())
```





